draft 1

### 2016-06-30 finalwriteup/draft.rmd
started conclusion

library('knitr')
library('ggplot2')
library("plyr")
library("dplyr")
library('magrittr')
library('reshape2')
library("rpart")
library('car')
library('e1071')
library('party')
library('randomForest')
library('RColorBrewer')
library('pwr')

@@

The project outcomes are intended to assist the business in  predicting project costs before project engagement and other business analyses such as client analysis. Several statistical and machine learning techniques will be tested, compared and refined.

@@

results.A2 = readRDS("report_data/reg_lm.rds")
results.A2 <- dplyr::filter(results.A2, diff>-15)
# n.samples(data = results.A2$diff)

@@

```{r load_rf_results}
results.rf.core = readRDS("report_data/reg_rf_core.rds")
# n.samples(data = results.rf.core$diff)

results.rf2 = readRDS("report_data/reg_rf.rds")
# n.samples(data = results.rf2$diff)

@@

```{r, fig.height=8}

@@

# How to achieve statistical power of 0.8?
finance.plot = readRDS("report_data/plot_finance.rds" )
fin.summary.1 = finance.plot %>% group_by(threshold, method) %>% 
        summarise_each(funs(mean_profit_ratio = round(mean(.), 1), 
                            sd_prof_ratio = round(sd(.), 1)),
                       profit.ratio)

@@

finance.plot = readRDS("report_data/plot_finance_100_4.rds" )
finance.plot$method <- factor(finance.plot$method, 
                                    levels=c("average", "simp.log","simp.boost",
                                             "comp.boost","comp.rf", "FWLS","orig.log",
                                             "orig.rf","orig.boost"))

@@

kable(best.fin)
# saveRDS(best.fin, "report_data/best_fin.rds")

@@

```{r TP_plot}
conf.plot <- readRDS("report_data/conf_plot.rds")

@@

conf.plot <- readRDS("report_data/conf_plot.rds")

@@

```{r}
reject.plot <- readRDS('report_data/reject_plot.rds')

@@

The aim of this project was to predict the profitability of projects for consulting businesses using their internal CRM data. Specifically, a statistical or machine learning model was created based on historical project data. The second aim was to test whether the model would have a positive impact on the overall profitability of a case study business. These research aims were fulfilled and the prediction model was refined to a binary classification problem of whether a project would be profitable or loss-making. This suited the limited data size best which was also highly complex. After testing several machine-learning and statistical models, a blended model of logistic regression, random forests, and boosted tree models predicted project profitability best. When applied to improving overall profitability of the business, the blended model which simply averaged the raw probability results from the three aforementioned models outperformed the individual models as well as the logistic regression blended  model. Final results showed the model could improve profitability of the business by between `r (-1.96*best.sd + best.mean) %>% round(1)`% and `r (1.96*best.sd + best.mean) %>% round(1)`% (the 95% confidence interval) by rejecting projects that have a 'probability' of being a loss-making job over `r best.t` (approximately `r ((simp.l %>% filter(mean.iter>best.t) %>% nrow)/nrow(simp.l)*100) %>% round(1)`% of projects).

>>>>

library('knitr')
library('ggplot2')
library("plyr")
library("dplyr")
library('magrittr')
library('reshape2')
library("rpart")
library('car')
library('e1071')
library('party')
library('randomForest')
library('RColorBrewer')
library('pwr')
library('scales')

@@

The project outcomes are intended to assist the business in  predicting project costs before project engagement. Several statistical and machine learning techniques will be tested, compared and refined.

@@

results.A2 = readRDS("reg_lm.rds")
results.A2 <- dplyr::filter(results.A2, diff>-15)
# n.samples(data = results.A2$diff)

@@

```{r load_rf_results}
results.rf.core = readRDS("reg_rf_core.rds")
# n.samples(data = results.rf.core$diff)

results.rf2 = readRDS("reg_rf.rds")
# n.samples(data = results.rf2$diff)

@@

```{r blend_violin, fig.height=8}

@@

# How to achieve statistical power of 0.8?
finance.plot = readRDS("plot_finance.rds" )
fin.summary.1 = finance.plot %>% group_by(threshold, method) %>% 
        summarise_each(funs(mean_profit_ratio = round(mean(.), 1), 
                            sd_prof_ratio = round(sd(.), 1)),
                       profit.ratio)

@@

finance.plot = readRDS("plot_finance_100_4.rds" )
finance.plot$method <- factor(finance.plot$method, 
                                    levels=c("average", "simp.log","simp.boost",
                                             "comp.boost","comp.rf", "FWLS","orig.log",
                                             "orig.rf","orig.boost"))

@@

kable(best.fin)
# saveRDS(best.fin, "best_fin.rds")

@@

```{r TP_plot}
conf.plot <- readRDS("conf_plot.rds")

@@

conf.plot <- readRDS("conf_plot.rds")

@@

```{r}
reject.plot <- readRDS('reject_plot.rds')

@@

The project outcomes are intended to assist the business in  predicting project costs before project engagement. Several statistical and machine learning techniques will be tested, compared and refined.

The objectives of this research consisted of a general aim and two research hypotheses, all of which have been fulfilled and tested. The findings made a significant contribution to the present body of literature and highlighted areas of future work that would refine the outcomes.

The general aim was to use statistical techniques to model the profitability of projects for a case study consulting business using their internal CRM data. This was rigorously completed by approaching the prediction of 'profitability' in two ways: as a regression problem predicting the degree of profitablility and as a binary classification problem predicting simply profit or loss. Several statistical and machine learning approaches were applied to both problems including naive bayes, bayesian networks, linear regression, random forests, gradient boosted trees and multiple methods of blending the individual models.

Hypothesis 1 stated a statistical or machine learning model based on the CRM data could predict the profitability of a new project. Testing this hypothesis revealed that the degree of profitability (the regression problem) could not be accurately predicted but whether a job would be profitable or not (binary classification) could be predicted to a level significantly better than random assignment (AUC = 0.76). This was achieved by 3 individual methods while various techniques that blended the individual methods improved results further (AUC = 0.77). It is likely the simpler task of binary classification suited the complex yet small data set better.

Lastly, hypothesis 2 stated the best predictive model developed from hypothesis 1 could be shown to improve the overall profitability (bottom l ine) of the case study business. This theory was tested on a range of predictive models because their AUC scores were not directly linked to profitability once the models were applied to a business scenario.The models output a probability, so a business scenario was tested where a probability threshold was chosen, and if a project was scored below the threshold it was accepted while projects above the threshold were rejected. If a project was accepted, the profits and losses were summed while if a project was rejected the profits were forfeited. If a model performed perfectly, all losses

After testing several machine-learning and statistical models, a blended model of logistic regression, random forests, and boosted tree models predicted project profitability best. When applied to improving overall profitability of the business, the blended model which simply averaged the raw probability results from the three aforementioned models outperformed the individual models as well as the logistic regression blended  model. Final results showed the model could improve profitability of the business by between `r (-1.96*best.sd + best.mean) %>% round(1)`% and `r (1.96*best.sd + best.mean) %>% round(1)`% (the 95% confidence interval) by rejecting projects that have a 'probability' of being a loss-making job over `r best.t` (approximately `r ((simp.l %>% filter(mean.iter>best.t) %>% nrow)/nrow(simp.l)*100) %>% round(1)`% of projects).
